Variational Inference for Inverse Reinforcement Learning with Gaussian Processes

### 2019-03-23 mpaper/paper.tex
It's working (I think)!

\section{Theoretical Justification} \label{sec:proof}
% TODO: Possible improvements:
% 1. Use limits more often.
% 2. Perturbation lemma can be replaced by continuity of A |-> A^-1.

@@

    \draw [->] (perturbation) edge (lemma3);
    \draw [->] (perturbation) edge (lemma1);
    \draw [->] (lebesgue) edge (theorem);
    \draw [->] (derivatives) edge (lemma1);
    \draw [->] (derivatives) edge (lemma2);
    \draw [->] (derivatives) edge (lemma3);
    \draw [->] (measurability) edge (theorem);
    \draw [->] (boundedness) edge (theorem);
    \draw [->] (lemma1) edge (theorem);
    \draw [->] (lemma2) edge (theorem);
    \draw [->] (lemma3) edge (theorem);
    \draw [->] (lemma4) edge (theorem);
  \end{tikzpicture}
  \caption{A graphical representation of dependencies between our theoretical
    results. An arrow from $A$ to $B$ means that $A$ was used to prove $B$.
    Results from the literature are in gray.}
  \label{fig:theorems}
\end{figure}

The typical way to optimise a quantity (the ELBO, in this case) involves
computing its gradient. Unfortunately, the term $\mathbb{E}[v]$ in $\mathcal{L}$
complicates the situation. The goal of this section is to show how Theorem
\ref{thm:lebesgue} can be applied to our model in order to derive the gradient
anyway\footnote{This technique is inspired by black box VI
  \cite{DBLP:conf/aistats/RanganathGB14}, but takes a more detailed look at the
  problem and requires significantly more work to prove correctness.}. After
showing that the theorem applies to our situation, we can estimate
$\nabla\mathbb{E}[v]$ with
\begin{align*}
  \nabla\mathbb{E}[v] &= \nabla \iint q(\mathbf{r} \mid \mathbf{u}) q(\mathbf{u})\dx \\
                      &= \iint \nabla[v q(\mathbf{r} \mid \mathbf{u}) q(\mathbf{u})]\dx \\
                      &= \iint \frac{\nabla[v q(\mathbf{r} \mid \mathbf{u})q(\mathbf{u})]}{q(\mathbf{r} \mid \mathbf{u})q(\mathbf{u})} q(\mathbf{r} \mid \mathbf{u}) q(\mathbf{u})\dx \\
                      &\approx \frac{1}{S} \sum_{s=1}^S \frac{\nabla[v q(\mathbf{r}_s \mid \mathbf{u}_s)q(\mathbf{u}_s)]}{q(\mathbf{r}_s \mid \mathbf{u}_s)q(\mathbf{u}_s)},
\end{align*}
which can be computed by drawing $S$ samples $\{(\mathbf{u}_s,
\mathbf{r}_s)\}_{s=1}^S$ from $\approximation$.

Our main goal is Theorem \ref{thm:main}, which allows us to move differentiation
inside the integral. In order to prove it, we use a number of intermediate
results. We start by stating a few derivatives of PDFs and covariance matrices,
and bound their values with some easy-to-deal-with polynomials. We then provide
a sketch proof of the measurability of MDP value functions, which is non-obvious
due to their non-trivial definition. Afterwards, we establish bounds for the
value functions, and, after another quick lemma, tackle the main proof of this
paper. See Figure \ref{fig:theorems} for an overview of how these results fit
together.

@@

% TODO: too long, lemma could be replaced by the continuity argument (and arrow
% removed from the figure), needs more explanations. Otherwise quite tolerable.
\begin{lemma} \label{lemma:bound1}
  Let $i \in \{ 0, \dots, d \}$ and $\epsilon > 0$ be arbitrary. Furthermore,
  let $c : \mathbb{R}^{|\mathcal{S}|} \times \mathbb{R}^m \to (\lambda_i -
  \epsilon, \lambda_i + \epsilon) \subset \mathbb{R}$ be a function with a
  codomain arbitrarily close to $\lambda_i$. Then
  \[
    \left. \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial \lambda_i}
    \right|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}
  \]
  has upper and lower bounds of the form $q(\mathbf{r} \mid
  \mathbf{u})d(\mathbf{u})$, where $d(\mathbf{u}) \in \mathbb{R}_2[\mathbf{u}]$.
\end{lemma}
\begin{proof}
  Remember that
  \begin{align*}
    \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial \lambda_i} &= \frac{1}{2}q(\mathbf{r} \mid \mathbf{u}) (|\bm\Gamma|^{-1} \tr(\mathbf{R} \adj(\bm\Gamma)) \\
                                                                      &- (\mathbf{r} - \mathbf{Su})^\intercal\bm\Gamma^{-1}\mathbf{R}\bm\Gamma^{-1}(\mathbf{r} - \mathbf{Su})).
  \end{align*}
  by Lemma \ref{lemma:derivatives}. Let $\mathbf{K}$ be any covariance
  matrix and
  \[
    \mathbf{A} = \frac{1}{\lambda_0} \mathbf{K}.
  \]

  % K and its derivative
  First, we can easily deduce that
  \begin{equation} \label{eq:convergence1}
    \mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} \to \mathbf{K}
  \end{equation}
  and
  \begin{equation} \label{eq:convergence2}
    \left. \frac{\partial \mathbf{K}}{\partial \lambda_i} \right|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} \to \frac{\partial \mathbf{K}}{\partial \lambda_i}
  \end{equation}
  as $\epsilon \to 0$. For \eqref{eq:convergence1}, the result is obvious if $i
  = 0$. Otherwise, it follows from the continuity of the exponential function.
  For \eqref{eq:convergence2}, observe that if $i=0$, then
  \[
    \left. \frac{\partial \mathbf{K}}{\partial \lambda_0} \right|_{\lambda_0 =
      c(\mathbf{r}, \mathbf{u})}
  \]
  as an expression has no $\lambda_0$ in it, so
  \[
    \left. \frac{\partial \mathbf{K}}{\partial \lambda_0} \right|_{\lambda_0 =
      c(\mathbf{r}, \mathbf{u})} = \frac{\partial \mathbf{K}}{\partial
      \lambda_0}.
  \]
  Finally, if $i>0$, then each element of 
  \[
    \left. \frac{\partial \mathbf{K}}{\partial \lambda_i}
    \right|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}
  \]
  is a constant multiple of the corresponding element of
  \[
    \mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})},
  \]
  so the same reasoning as for \eqref{eq:convergence1} applies.

  % Inverse of a covariance matrix
  Now, we will show that $\mathbf{K}^{-1}|_{\lambda_i = c(\mathbf{r},
    \mathbf{u})}$ exists and
  \[
    \lim_{\epsilon \to 0} \mathbf{K}^{-1}|_{\lambda_i = c(\mathbf{r},
      \mathbf{u})} = \mathbf{K}.
  \]
  If $i = 0$, then $\mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} =
  c(\mathbf{r}, \mathbf{u})\mathbf{A}$. Therefore\footnote{Note that since
    $\lambda_0 \ne 0$, $c(\mathbf{r}, \mathbf{u}) \ne 0$ for small enough
    $\epsilon$.},
  \begin{align*}
    \mathbf{K}^{-1}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} &=
                                                               \frac{1}{c(\mathbf{r}, \mathbf{u})}\mathbf{A}^{-1} \\
                                                             &\to \frac{1}{\lambda_0}\mathbf{A}^{-1} = \frac{1}{\lambda_0} \left( \frac{1}{\lambda_0}\mathbf{K} \right)^{-1} = \mathbf{K}^{-1}
  \end{align*}
  as $\epsilon \to 0$. For $i > 0$, let
  \[
    S = \sum_{n \in \{ 1, \dots, d \} \setminus \{ i \}} \frac{\lambda_n}{2}
    (x_{j,n} - x_{k,n})^2 + \mathbbm{1}[j \ne k]\sigma^2 \lambda_n
  \]
  and $\delta = c(\mathbf{r}, \mathbf{u}) - \lambda_i$ so that $c(\mathbf{r},
  \mathbf{u}) = \lambda_i + \delta$, and $\lim_{\epsilon \to 0} \delta = 0$.
  Then,
  \begin{align*}
    &k(\mathbf{x}_j, \mathbf{x}_k)|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} = \!\begin{multlined}[t]
      \lambda_0 \exp \left( \vphantom{\frac{1}{2}} \right. -\frac{1}{2} c(\mathbf{r}, \mathbf{u})(x_{j,i} - x_{k,i})^2\\
      - \mathbbm{1}[j \ne k]\sigma^2c(\mathbf{r}, \mathbf{u}) - S \left. \vphantom{\frac{1}{2}} \right)
    \end{multlined} \\
    &= \!\begin{multlined}[t]
      \lambda_0 \exp \left( \vphantom{\frac{1}{2}} \right. -\frac{1}{2} (\lambda_i + \delta)(x_{j,i} - x_{k,i})^2 \\
      - \mathbbm{1}[j \ne k]\sigma^2(\lambda_i + \delta) - S \left. \vphantom{\frac{1}{2}} \right)
    \end{multlined} \\
    &= \!\begin{multlined}[t]
      \lambda_0 \exp \left( \vphantom{\frac{1}{2}} \right. -\frac{1}{2}(\mathbf{x}_j - \mathbf{x}_k)^\intercal \bm\Lambda (\mathbf{x}_j - \mathbf{x}_k) - \mathbbm{1}[j \ne k]\sigma^2\tr(\bm\Lambda) \\
      - \frac{\delta}{2}(x_{j,i} - x_{k,i})^2 - \mathbbm{1}[j \ne k]\sigma^2\delta \left. \vphantom{\frac{1}{2}} \right)
    \end{multlined} \\
    &= k(\mathbf{x}_j, \mathbf{x}_k)\exp \left( \vphantom{\frac{1}{2}} \right. -\frac{\delta}{2}(x_{j,i} - x_{k,i})^2 - \mathbbm{1}[j \ne k]\sigma^2\delta \left. \vphantom{\frac{1}{2}} \right) \\
    &= \!\begin{multlined}[t]
      k(\mathbf{x}_j, \mathbf{x}_k) + k(\mathbf{x}_j, \mathbf{x}_k) \left( \exp
        \left( -\frac{\delta}{2}(x_{j,i} - x_{k,i})^2 \right. \right. \\
      -\mathbbm{1}[j \ne k]\sigma^2\delta \left. \left. \vphantom{\frac{1}{2}} \right) - 1 \right).
    \end{multlined}
  \end{align*}
  Thus, we can express $\mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$ as
  $\mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} = \mathbf{K} +
  \mathbf{E}$, where $\mathbf{E}$ is defined by
  \[
    E_{j,k} = k(\mathbf{x}_j, \mathbf{x}_k) \left( \exp \left(
        -\frac{\delta}{2}(x_{j,i} - x_{k,i})^2 \right. \right. -\mathbbm{1}[j
    \ne k]\sigma^2\delta \left. \left. \vphantom{\frac{1}{2}} \right) - 1
    \right).
  \]
  By this definition,
  \[
    \lim_{\epsilon \to 0} E_{j,k} = 0.
  \]
  Then, since $\mathbf{K}$ is invertible,
  \[
    \mathbf{K}^{-1}_{\lambda_i = c(\mathbf{r}, \mathbf{u})} \to \mathbf{K}^{-1}
  \]
  by continuity of $\mathbf{A} \mapsto \mathbf{A}^{-1}$.

  This is enough to prove constant upper and lower bounds on
  $\mathbf{S}$, $\bm\Gamma$, and $\mathbf{R}$ (all with $\lambda_i$ replaced
  with $c(\mathbf{r}, \mathbf{u})$), which means that $(\mathbf{r} -
  \mathbf{Su})^\intercal\bm\Gamma^{-1}\mathbf{R}\bm\Gamma^{-1}(\mathbf{r} -
  \mathbf{Su})|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$ has upper and lower
  bounds in $\mathbb{R}_2[\mathbf{u}]$. Furthermore, having convergence results
  for arbitrary covariance matrices and their inverses means that
  \[
    \lim_{\epsilon \to 0} \bm\Gamma|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} =
    \bm\Gamma.
  \]
  Hence,
  \[
    \lim_{\epsilon \to 0} \det(\bm\Gamma)|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} =
    \det(\bm\Gamma).
  \]
  Assuming that
  $\bm\Gamma$ is invertible so that $q(\mathbf{r} \mid \mathbf{u})$ exists,
  \[
    \det(\bm\Gamma)|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} \ne 0
  \]
  for small enough $\epsilon$, and, thus, $\det(\bm\Gamma)^{-1}|_{\lambda_i =
    c(\mathbf{r}, \mathbf{u})}$ exists and is bounded.

  Recall that
  \[
    \bm\Gamma = \Krr - \mathbf{S}\Kru = \Krr - \Kru^\intercal\Kuu^{-1}\Kru.
  \]
  We have already demonstrated that for $\mathbf{K} \in \{ \Krr, \Kru, \Kuu^{-1}
  \}$,
  \[
    \lim_{\epsilon \to 0}\mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} =
    \mathbf{K}.
  \]
  We can use a general fact that if $\lim_{x \to 0} f(x) = f$,
  then $f(x) = f + g(x)$ for some function $g$ such that $\lim_{x \to 0} g(x) =
  0$ to define matrices $\Err$, $\Eru$, and $\Euu$ such that
  \begin{alignat*}{3}
    \Krr|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} &= \Krr + \Err, &&\Err \to
    \mathbf{O}_{|\mathcal{S}|}, \\
    \Kru|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} &= \Kru + \Eru, \quad
    \text{and} \quad &&\Eru \to \mathbf{O}_{|\mathcal{S}|, m}, \\
    \Kuu^{-1}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} &= \Kuu^{-1} + \Euu,
    &&\Euu \to \mathbf{O}_{m}.
  \end{alignat*}
  as $\epsilon \to 0$. Then, $\bm\Gamma|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}
  = \bm\Gamma + \mathbf{E}$, where
  \begin{align*}
    \mathbf{E} &= \Err - \Kru^\intercal\Kuu^{-1}\Eru - \Kru^\intercal\Euu(\Kru + \Eru) \\
               &- \Eru^\intercal(\Kuu^{-1} + \Euu)(\Kru + \Eru) \\
               &\to \mathbf{O}_{|\mathcal{S}|}
  \end{align*}
  as $\epsilon \to 0$. Thus, Lemma \ref{lemma:perturbation} shows that
  $\bm\Gamma^{-1}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$ exists and provides
  constant bounds on its elements.

  Since we already know that $\bm\Gamma^{-1}|_{\lambda_i = c(\mathbf{r},
    \mathbf{u})}$ and $\det(\bm\Gamma)|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$
  are bounded, so is $\adj(\bm\Gamma) =
  \det(\bm\Gamma)\bm\Gamma^{-1}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$. Thus,
  we have constant bounds on
  $|\bm\Gamma|^{-1}\tr(\mathbf{R}\adj(\bm\Gamma))|_{\lambda_i = c(\mathbf{r},
    \mathbf{u})}$, which means that
  \[
    \left. \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial \lambda_i}
    \right|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}
  \]
  has the required bounds.
\end{proof}

@@

Our main theorem is a specialised version of an integral differentiation result
by Timoney \cite{lecture_notes}.
\begin{theorem} \label{thm:main}
  Whenever the derivative exists,
  \[
    \dt\iint
    \V(s)q(\mathbf{r} \mid \mathbf{u})q(\mathbf{u})\dx
    = \iint
    \dt[\V(s)q(\mathbf{r} \mid \mathbf{u})q(\mathbf{u})]\dx,
  \]
  where $t$ is any scalar part of $\bm\mu$, $\bm\Sigma$, or $\bm\lambda$.
\end{theorem}
\begin{proof}
  Let
  \begin{align*}
    \f &= \V(s)q(\mathbf{r} \mid \mathbf{u})q(\mathbf{u}), \\
    F(t) &= \iint \f\dx,
  \end{align*}
  and fix the value of $t$. Let $(t_n)_{n=1}^\infty$ be any sequence such that
  $\lim_{n \to \infty} t_n = t$, but $t_n \ne t$ for all $n$. We want to show
  that
  \begin{equation} \label{eq:to_prove}
    F'(t) = \lim_{n \to \infty} \frac{F(t_n) - F(t)}{t_n - t} = \iint \df\dx.
  \end{equation}
  We have
  \[
    \begin{split}
      \frac{F(t_n) - F(t)}{t_n - t} &= \iint \frac{\ftn - \f}{t_n - t}\dx \\
      &= \iint \fn\dx,
    \end{split}
  \]
  where
  \[
    \fn = \frac{\ftn - \f}{t_n - t}.
  \]
  Since
  \[
    \lim_{n \to \infty} \fn = \df,
  \]
  \eqref{eq:to_prove} follows from Theorem \ref{thm:lebesgue} as soon as we show
  that both $f$ and $f_n$ are measurable and find a non-negative integrable
  function $g$ such that for all $n$, $\mathbf{r}$, $\mathbf{u}$,
  \[
    |\fn| \le \g.
  \]
  The MDP value function is measurable by Proposition \ref{thm:measurability}.
  The result of multiplying or adding measurable functions (e.g., probability
  density functions) to a measurable function is still measurable. Thus,
  both $f$ and $f_n$ are measurable.

  It remains to find $g$. For notational simplicity and without loss of
  generality, we will temporarily assume that $t$ is a parameter of
  $q(\mathbf{r} \mid \mathbf{u})$. Then
  \[
    |\fn| = |\V(s)| \left| \frac{q(\mathbf{r} \mid \mathbf{u})|_{t =
          t_n} - q(\mathbf{r} \mid \mathbf{u})}{t_n - t} \right| q(\mathbf{u})
  \]
  since PDFs are non-negative. An upper bound for
  $|\V(s)|$ is given by Proposition \ref{thm:bound}, while
  \[
    \frac{q(\mathbf{r} \mid \mathbf{u})|_{t = t_n} - q(\mathbf{r} \mid \mathbf{u})}{t_n - t} = \left.
      \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial t} \right|_{t = c(\mathbf{r},
      \mathbf{u})}
  \]
  for some function $c : \mathbb{R}^{|\mathcal{S}|} \times \mathbb{R}^m \to
  (\min\{t, t_n\}, \max\{t, t_n\})$ due to the mean value theorem (since $q$ is
  a continuous and differentiable function of $t$, regardless of the specific
  choices of $q$ and $t$).

>>>>

\section{Theoretical Justification} \label{sec:proof}
% TODO: Possible improvements:
% 1. Use limits more often.
% 2. Perturbation lemma can be replaced by continuity of A |-> A^-1.
% 3. epsilon is treated as an arbitrary constant.

The typical way to optimise a quantity (the ELBO, in this case) involves
computing its gradient. Unfortunately, the term $\mathbb{E}[v]$ in $\mathcal{L}$
complicates the situation. The goal of this section is to show how Theorem
\ref{thm:lebesgue} can be applied to our model in order to derive the gradient
anyway\footnote{This technique is inspired by black box VI
  \cite{DBLP:conf/aistats/RanganathGB14}, but takes a more detailed look at the
  problem and requires significantly more work to prove correctness.}. After
showing that the theorem applies to our situation, we can estimate
$\nabla\mathbb{E}[v]$ with
\begin{align*}
  \nabla\mathbb{E}[v] &= \nabla \iint v q(\mathbf{r} \mid \mathbf{u}) q(\mathbf{u})\dx \\
                      &= \iint \nabla[v q(\mathbf{r} \mid \mathbf{u}) q(\mathbf{u})]\dx \\
                      &= \iint \frac{\nabla[v q(\mathbf{r} \mid \mathbf{u})q(\mathbf{u})]}{q(\mathbf{r} \mid \mathbf{u})q(\mathbf{u})} q(\mathbf{r} \mid \mathbf{u}) q(\mathbf{u})\dx \\
                      &\approx \frac{1}{S} \sum_{s=1}^S \frac{\nabla[v q(\mathbf{r}_s \mid \mathbf{u}_s)q(\mathbf{u}_s)]}{q(\mathbf{r}_s \mid \mathbf{u}_s)q(\mathbf{u}_s)},
\end{align*}
which can be computed by drawing $S$ samples $\{(\mathbf{u}_s,
\mathbf{r}_s)\}_{s=1}^S$ from $\approximation$.

@@

    \draw [->] (perturbation) edge (lemma3);
    \draw [->] (lebesgue) edge (theorem);
    \draw [->] (derivatives) edge (lemma1);
    \draw [->] (derivatives) edge (lemma2);
    \draw [->] (derivatives) edge (lemma3);
    \draw [->] (measurability) edge (theorem);
    \draw [->] (boundedness) edge (theorem);
    \draw [->] (lemma1) edge (theorem);
    \draw [->] (lemma2) edge (theorem);
    \draw [->] (lemma3) edge (theorem);
    \draw [->] (lemma4) edge (theorem);
  \end{tikzpicture}
  \caption{A graphical representation of dependencies between our theoretical
    results. An arrow from $A$ to $B$ means that $A$ was used to prove $B$.
    Results from the literature are in gray.}
  \label{fig:theorems}
\end{figure}

Our main goal is Theorem \ref{thm:main}, which allows us to move differentiation
inside the integral. In order to prove it, we use a number of intermediate
results. We start by stating a few derivatives of PDFs and covariance matrices,
and bound their values with some easy-to-deal-with polynomials. We then provide
a sketch proof of the measurability of MDP value functions as well as new upper
and lower bounds on their values. After another quick lemma, we tackle the main
proof of this paper. See Figure \ref{fig:theorems} for an overview of how these
results fit together.

@@

\begin{lemma} \label{lemma:bound1}
  Let $i \in \{ 0, \dots, d \}$ be arbitrary, and let $c :
  \mathbb{R}^{|\mathcal{S}|} \times \mathbb{R}^m \to (\lambda_i - \epsilon,
  \lambda_i + \epsilon) \subset \mathbb{R}$ be a function with a codomain
  arbitrarily close to $\lambda_i$. Then
  \[
    \left. \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial \lambda_i}
    \right|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}
  \]
  has upper and lower bounds of the form $q(\mathbf{r} \mid
  \mathbf{u})d(\mathbf{u})$, where $d(\mathbf{u}) \in \mathbb{R}_2[\mathbf{u}]$.
\end{lemma}
\begin{proof}
  Let $\mathbf{K}$ be any covariance matrix and set
  \[
    \mathbf{A} = \frac{1}{\lambda_0} \mathbf{K}.
  \]

  % K and its derivative
  First, we can easily deduce that\footnote{Note that we often switch between
    proving limits and bounds. We are fundamentally interested in proving
    constant bounds, but sometimes it is more convenient to show convergence
    instead. In this situation, convergence as $\epsilon \to 0$ implies constant
    upper and lower bounds.}
  \begin{equation} \label{eq:convergence1}
    \mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} \to \mathbf{K}
  \end{equation}
  and
  \begin{equation} \label{eq:convergence2}
    \left. \frac{\partial \mathbf{K}}{\partial \lambda_i} \right|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} \to \frac{\partial \mathbf{K}}{\partial \lambda_i}
  \end{equation}
  as $\epsilon \to 0$. For \eqref{eq:convergence1}, the result is obvious if $i
  = 0$. Otherwise, it follows from the continuity of the exponential function.
  For \eqref{eq:convergence2}, observe that if $i=0$, then
  \[
    \left. \frac{\partial \mathbf{K}}{\partial \lambda_0} \right|_{\lambda_0 =
      c(\mathbf{r}, \mathbf{u})}
  \]
  as an expression contains no $\lambda_0$, so
  \[
    \left. \frac{\partial \mathbf{K}}{\partial \lambda_0} \right|_{\lambda_0 =
      c(\mathbf{r}, \mathbf{u})} = \frac{\partial \mathbf{K}}{\partial
      \lambda_0}.
  \]
  Finally, if $i>0$, then each element of 
  \[
    \left. \frac{\partial \mathbf{K}}{\partial \lambda_i}
    \right|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}
  \]
  is a constant multiple of the corresponding element of
  \[
    \mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})},
  \]
  so the same reasoning applies as in case of \eqref{eq:convergence1}.

  % Inverse of a covariance matrix
  Next, we will show that $\mathbf{K}^{-1}|_{\lambda_i = c(\mathbf{r},
    \mathbf{u})}$ exists and
  \begin{equation} \label{eq:limit3}
    \lim_{\epsilon \to 0} \mathbf{K}^{-1}|_{\lambda_i = c(\mathbf{r},
      \mathbf{u})} = \mathbf{K}^{-1}.
  \end{equation}
  If $i = 0$, then $\mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} =
  c(\mathbf{r}, \mathbf{u})\mathbf{A}$. Therefore\footnote{Note that since
    $\lambda_0 > 0$, $c(\mathbf{r}, \mathbf{u}) \ne 0$ for small enough
    $\epsilon$.},
  \begin{align*}
    \mathbf{K}^{-1}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} &=
                                                               \frac{1}{c(\mathbf{r}, \mathbf{u})}\mathbf{A}^{-1} \\
                                                             &\to \frac{1}{\lambda_0}\mathbf{A}^{-1} = \frac{1}{\lambda_0} \left( \frac{1}{\lambda_0}\mathbf{K} \right)^{-1} = \mathbf{K}^{-1}
  \end{align*}
  as $\epsilon \to 0$. For $i > 0$, by \eqref{eq:convergence1} and continuity of
  $\mathbf{A} \mapsto \mathbf{A}^{-1}$ we immediate get \eqref{eq:limit3}.

  This is enough to prove constant upper and lower bounds on
  $\mathbf{S}$, $\bm\Gamma$, and $\mathbf{R}$ (all with $\lambda_i$ replaced
  with $c(\mathbf{r}, \mathbf{u})$), which means that $(\mathbf{r} -
  \mathbf{Su})^\intercal\bm\Gamma^{-1}\mathbf{R}\bm\Gamma^{-1}(\mathbf{r} -
  \mathbf{Su})|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$ has upper and lower
  bounds in $\mathbb{R}_2[\mathbf{u}]$. Furthermore, having convergence results
  for arbitrary covariance matrices and their inverses means that
  \begin{equation} \label{eq:gamma_convergence}
    \lim_{\epsilon \to 0} \bm\Gamma|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} =
    \bm\Gamma
  \end{equation}
  and, by continuity of the determinant function,
  \begin{equation} \label{eq:det_gamma}
    \lim_{\epsilon \to 0} \det(\bm\Gamma)|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} =
    \det(\bm\Gamma).
  \end{equation}
  Assuming that
  $\bm\Gamma$ is invertible so that $q(\mathbf{r} \mid \mathbf{u})$ exists,
  \[
    \det(\bm\Gamma)|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} \ne 0
  \]
  for small enough $\epsilon$, and, thus, $\det(\bm\Gamma)^{-1}|_{\lambda_i =
    c(\mathbf{r}, \mathbf{u})}$ exists and is bounded.

  From \eqref{eq:gamma_convergence} and continuity of $\mathbf{A} \mapsto
  \mathbf{A}^{-1}$ we can immediately deduce that
  \begin{equation} \label{eq:gamma_inverse}
    \lim_{\epsilon \to 0} \bm\Gamma^{-1}|_{\lambda_i = c(\mathbf{r},
      \mathbf{u})} = \bm\Gamma^{-1}.
  \end{equation}
  Similarly, from \eqref{eq:det_gamma} and \eqref{eq:gamma_inverse} we get that
  \[
    \adj(\bm\Gamma)|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} =
    \det(\bm\Gamma)\bm\Gamma^{-1}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} \to
    \det(\bm\Gamma)\bm\Gamma^{-1}
  \]
  as $\epsilon \to 0$. This gives us constant bounds on
  \[
    |\bm\Gamma|^{-1}\tr(\mathbf{R}\adj(\bm\Gamma))|_{\lambda_i = c(\mathbf{r},
      \mathbf{u})},
  \]
  which completes the proof that
  \[
    \left. \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial \lambda_i}
    \right|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}
  \]
  has the required quadratic bounds.
\end{proof}

@@

Our main theorem is a specialised version of an integral differentiation result
by Timoney \cite{lecture_notes}.
\begin{theorem} \label{thm:main}
  Whenever the derivative exists,
  \[
    \dt\iint
    \V(s)q(\mathbf{r} \mid \mathbf{u})q(\mathbf{u})\dx
    = \iint
    \dt[\V(s)q(\mathbf{r} \mid \mathbf{u})q(\mathbf{u})]\dx,
  \]
  where $t$ is any scalar part of $\bm\mu$, $\bm\Sigma$, or $\bm\lambda$.
\end{theorem}
\begin{proof}
  Let
  \begin{align*}
    \f &= \V(s)q(\mathbf{r} \mid \mathbf{u})q(\mathbf{u}), \\
    F(t) &= \iint \f\dx,
  \end{align*}
  and fix the value of $t$. Let $(t_n)_{n=1}^\infty$ be any sequence such that
  $\lim_{n \to \infty} t_n = t$, but $t_n \ne t$ for all $n$. We want to show
  that
  \begin{equation} \label{eq:to_prove}
    F'(t) = \lim_{n \to \infty} \frac{F(t_n) - F(t)}{t_n - t} = \iint \df\dx.
  \end{equation}
  We have
  \[
    \begin{split}
      \frac{F(t_n) - F(t)}{t_n - t} &= \iint \frac{\ftn - \f}{t_n - t}\dx \\
      &= \iint \fn\dx,
    \end{split}
  \]
  where
  \[
    \fn = \frac{\ftn - \f}{t_n - t}.
  \]
  Since
  \[
    \lim_{n \to \infty} \fn = \df,
  \]
  \eqref{eq:to_prove} follows from Theorem \ref{thm:lebesgue} as soon as we show
  that both $f$ and $f_n$ are measurable and find a non-negative integrable
  function $g$ such that for all $n$, $\mathbf{r}$, $\mathbf{u}$,
  \[
    |\fn| \le \g.
  \]
  The MDP value function is measurable by Proposition \ref{thm:measurability}.
  The result of multiplying or adding measurable functions (e.g., PDFs) to a
  measurable function is still measurable. Thus, both $f$ and $f_n$ are
  measurable.

  It remains to find $g$. For notational simplicity and without loss of
  generality, we will temporarily assume that $t$ is a parameter of
  $q(\mathbf{r} \mid \mathbf{u})$. Then
  \[
    |\fn| = |\V(s)| \left| \frac{q(\mathbf{r} \mid \mathbf{u})|_{t =
          t_n} - q(\mathbf{r} \mid \mathbf{u})}{t_n - t} \right| q(\mathbf{u}),
  \]
  since PDFs are non-negative. An upper bound for
  $|\V(s)|$ is given by Proposition \ref{thm:bound}, while
  \[
    \frac{q(\mathbf{r} \mid \mathbf{u})|_{t = t_n} - q(\mathbf{r} \mid \mathbf{u})}{t_n - t} = \left.
      \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial t} \right|_{t = c(\mathbf{r},
      \mathbf{u})}
  \]
  for some function $c : \mathbb{R}^{|\mathcal{S}|} \times \mathbb{R}^m \to
  (\min\{t, t_n\}, \max\{t, t_n\})$ due to the mean value theorem (since $q$ is
  a continuous and differentiable function of $t$, regardless of the specific
  choices of $q$ and $t$).
